### 2025-05-21 paper.qmd
Updating draft

---
title: Tail Flexibility in the Degrees of Preferential
 Attachment Networks
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-basic
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    number-sections: true
author:
  - name: Thomas William Boughen
    affiliations:
      - name: Newcastle University
        id: 1
        department: School of Mathematics, Statistics and Physics
  - name: Clement Lee
    affiliations:
      - ref: 1
  - name: Vianey Palacios Ramirez
    affiliations:
      - ref: 1
abstract: |
  Devising the underlying generating mechanism of a real-life network is difficult as, more often than not, only its snapshots are available, but not its full evolution. One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law. Consequently, the growth of real-life networks that roughly display such power-law behaviour is commonly modelled by preferential attachment. However, the validity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the right tail of the degree distribution is often lighter than implied by the body, whilst still being heavy. In this paper, we study a modified version of the model with a flexible preference function that allows super/sub-linear behaviour whilst also guaranteeing that the limiting degree distribution has a heavy tail. We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
keywords: [networks, extremes]
bibliography: refs.bib  
---

\newpage

```{r,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(paletteer)
library(gridExtra)
library(latex2exp)
library(ggpubr)
library(Rcpp)
library(RcppArmadillo)
```

# Introduction

Networks appear in many fields such as sociology, politics, epidemiology, and economics. Statistical methods have been and continue to be used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17].

One example of a mechanistic model that may be considered is the preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b(\cdot)$ of their in-degree. The well-known Barabási-Albert (BA) model is a special case of this when $b(k) = k+1$ [@Barabasi99].

These mechanistic models are generally useful when simulating networks, but not for fitting to real data as usually only snapshots of a network are available and not the full evolution. However, one statistic that is usually available for snapshots of networks is the degree distribution making it valuable to study the degrees in a network. Often, when looking at the degree distributions of real networks it can be attractive to use the power law when modelling as it is seemingly observed in much of the data. It also comes with the added benefit of offering a connection to preferential attachment since the BA model results in a power law degree distribution [@Barabasi99], perhaps suggesting that if the power law fits well then the network may have grown according to similar simple rules. However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as [@krapivsky01] shows that a general preference function does not lead to power law degrees and additionally, it is still being argued whether or not a lot of real data actually follows a power law. @Broido_2019 provides some evidence that suggests that real world networks rarely have degrees that follow a power law through the use of statistical analysis of almost 1000 networks across a wide variety of fields; @Voitalov_2019, on the other hand, disagrees and claims that these kinds of networks are not nearly as rare and only appear so as a result of an unrealistic expectation of power law without deviations or noise.

Extreme value methods have been scarcely used when it comes to analysis of degree distributions and whether they follow the power law. The absence of these methods essentially down weights the deviation of vertices with the largest degrees, which are usually the most influential, from the power law. Somewhat recent publications [@Voitalov_2019; @Lee24; @wang2022random] have addressed this gap in the literature. @Voitalov_2019 proposes estimating the power law exponent of a degree distribution using three estimators based in extreme value theory, in contrast to the estimators generally used in network science, the degree distributions are then split into one of three categories (not power-law , hardly power-law or power-law) depending on the values of the estimates obtained from each of the estimators. @Lee24 proposes a framework for modelling entire degree distributions using a combination of two generalisations of the power law including a model selection step that is capable of suggesting whether or not the data is adequately modelled by a power law, with one conclusion being that most data seems to have a tail lighter than what is implied by the body whilst still being heavy tailed. @wang2022random considers a preferential attachment model with an affine preference function and reciprocal edges, they study the limiting degree distribution of this model for both the in-degrees and out-degrees - finding that the joint limiting distribution is multivariate regularly varying and has the property of hidden regular variation.

Evidently, the study of degree distributions is a very active but the vast majority of research in the field (whether using extremes or not) are purely descriptive in the sense that no information about the preference function is revealed even if preferential attachment was the underlying mechanism. This paper aims to address this gap - asking if given the degree distribution of a network assumed to come from the preferential attachment model, can we directly infer the model parameters?

@rudas07 derives the limiting degree distribution in terms of the preference function $b(\cdot)$ and its parameters, we build upon this by finding a class of preference functions that result in realistic tails i.e. heavy but not as heavy as implied by the power law.

The remainder of this paper begins with a more detailed description of the PA model alongside the theoretical results for the limiting degree distribution using results from @rudas07, this section (@sec-tail) then goes on to analyse the tail behaviour of this limiting distribution in terms of the preference function $b(\cdot)$ arguing that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function - ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible. @sec-model continues on from the previous section by considering the limiting distribution obtained when using a preference function of the form stated at the end of @sec-tail providing examples of how the shape of the distribution varies with respect to the model parameters as well as visualisations for how the tail heaviness changes when these parameters are varied - continued study of this distribution then also reveals a connection to the general Pareto distribution (GPD). @sec-sim consists of a simulation study where networks are simulated from the PA model for various parameter combinations and then the model specified in @sec-model is fitted to the resulting degree distributions using Bayesian inference, demonstrating that (for simulated data) the model parameters can be recovered from only the degree distribution. Finally, @sec-real fits the model to some real data and provides posterior estimates for the preference function assuming the networks evolved according to the PA model.

\newpage

# Tail Behaviour of Preferential Attachment Model {#sec-tail}

The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:

Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree where the preference function $b$ is chosen such that:

$$
b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
$$ {#eq-condb1}

$$
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
$$ {#eq-condb2}

Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.

Given conditions [@eq-condb1] and [@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07]. Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:

$$
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
$$ where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.

The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k$, that is:

$$
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
$$ which by the previously stated theorem can also be written as:

$$
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$ Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:

$$
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
$$

We are specifically interested in how the tail heaviness ($\xi$ in the IGPD) is affected by the preference function $b$, [@shimura12] introduces a quantity that will help in determining what domain of attraction this discrete distribution belongs to, that is:

For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let:

$$
\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
$$

[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(k) \sim k^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light tailed. This allows us to show the following:

::: {#prp-omega}
If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then $$
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$ and

$$
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
$$
:::

See appendix [-@sec-proofs] for the details of the proof.

@prp-omega aligns with the result from [@krapivsky01] demonstrating that a sub-linear preference function will lead to a light tailed distribution. So, in order for the degree distribution to be heavy tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive. We can in fact show that BA model produces a heavy tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$, noting that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ is simply $\lim_{k\rightarrow\infty} 1 = 1$ leaving the tail heaviness to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.

We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e. $\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$. First we must consider the following theorem:

::: {#thm-stolz}
## Stolz-Cesàro Theorem [@cesaro]

Let $(a_n)_{n\ge1}$ and $(b_n)_{n\ge1}$ be two sequence of real numbers. Assume $(a_n)_{n\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:

$$
\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
$$ Then,

$$
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
$$
:::

From @prp-omega we know that in order for the limiting degree distribution to have heavy tails, we must have that:

$$
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
$$ Now, setting $b_k = b(k)$ and $a_k = k$:

$$
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
$$ meaning by @thm-stolz:

$$
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0.
$$ Therefore, in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear. With this under consideration we propose the following preference function:

$$
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.

This preference function, as per @prp-omega, will produce a degree distribution with tail heaviness $\beta/\lambda^*$ guaranteeing a heavy tail. We study this preference function further in the next section.

\newpage

# Preferential Attachment with Flexible Heavy Tail {#sec-model}

Using a preference function with guaranteed linear behaviour in the limit, allows for the inclusion of sub/super linear behaviour without losing the heavy tails or ending up with a degenerate degree distribution.

The limiting degree distribution resulting from using a preference function of this form can be found to have survival:

$$
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
\end{cases}
$$ {#eq-polysurv}

with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where:

$$
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
$$ {#eq-rho}

which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.

Some examples of what the degree distribution looks like are shown below in @fig-polylinsurv:

```{r, echo=FALSE}
source('scripts/funcs.R')
```

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical survival distributions of the limiting degree distributions, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
#| label: fig-polylinsurv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = 0:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
}
pars = cbind(pars,surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

ggplot(data = pars) + geom_line(aes(x=(x+1),y=surv, linetype=as.character(b),colour =as.character(b))) +
  scale_x_log10() + scale_y_log10(limits=c(1e-5,1))+theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) + 
  facet_grid(eps~a,labeller = labeller, scales='free')
```

The survival (@eq-polysurv) can be connected to the IGPD by using Stirling's approximation of the gamma function, where the IGPD is a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ defined by the survival: $$
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.

Using Stirling's approximation we have that: $$
\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
$$ now utilising the expression from @eq-polysurv:

\begin{align*}
\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
\end{align*}

$$
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
$$ meaning that for $k\ge k_0$ the limiting degree distribution is similar to $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.

To assess how close of an approximation this is the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey. The approximation seems to hold up fairly well even for large degrees.

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical conditional survivals (grey) alongside their IGPD approximations (coloured).
#| label: fig-approx_surv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = k0s:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
gp_surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])/
    S(pars$k0[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
  gp_surv[i] = mev::pgp(pars$x[i],pars$k0[i]-1,((pars$k0[i]+1)^pars$a[i] +pars$eps[i]-1 )/pars$lambda[i],pars$b[i]/pars$lambda[i],lower.tail = F)
}
pars = cbind(pars,surv,gp_surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv','gp_surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

library(ggh4x)

l1 = scale_y_log10(limits=c(1e-1,1))
l2 = scale_y_log10(limits=c(1e-2,1))
l3 = scale_y_log10(limits=c(1e-3,1))
l4 = scale_y_log10(limits=c(1e-4,1))

limlist = list(l1,l2,l3,l4)

ggplot(data = pars) + geom_line(aes(x=(x),y=surv, linetype=as.character(b),colour =as.character(b)), lwd=1) +
  geom_line(data = pars[pars$x>=pars$k0,],aes(x=(x+1),y=gp_surv, linetype=as.character(b)),colour ='grey')+
  scale_x_log10()+
  theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) +
  facet_grid(eps~a,labeller = labeller,scales='free') + facetted_pos_scales(y=limlist)

```

Since $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed. Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices:

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 12
#| fig-height: 6
#| fig-cap: Heat maps of $\xi$ for various combinations of the parameters of the proposed model.
#| label: fig-polyheat
#| 
rho_optim_ba= Vectorize(function(a,eps, b, k0){
  return(abs(rho(2*b, polylin(a, eps), b, k0)-1))
},vectorize.args = 'a')
find_a_ba = Vectorize(function(b, eps, k0){
  out = optimise(rho_optim_ba, c(0.00001,3), b=b, eps=eps, k0=k0)$minimum
  return(out)
}, vectorize.args = 'b')
N =50
as = seq(0,2,l=N+1)[-1]
bs = seq(0,2,l=N+1)[-1]
eps = c(.01,.1,.5,1)
k0 = c(25,100)
pars = expand.grid(as,bs,eps,k0)
names(pars) = c('a', 'b', 'eps', 'k0')
lambdas = numeric(nrow(pars))
a_for_ba = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
  a_for_ba[i] = find_a_ba(pars$b[i], pars$eps[i], pars$k0[i])
}
pars = cbind(pars,lambdas,a_for_ba)
names(pars) = c('a', 'b', 'eps', 'k0','lambda','ba')
labeller = label_bquote(cols = `epsilon`==.(eps),rows = ~k[0]==.(k0))

ggplot(pars) + geom_raster(aes(x=a,y=b,fill=b/lambda)) +
  geom_line(aes(x=ba, y=b),linetype='dashed', colour='red',lwd=1)+
  scale_fill_paletteer_c(palette='grDevices::Blues',limits=c(0,1),direction=-1)+theme(aspect.ratio = 1)+ylim(min(bs), max(bs))+xlim(min(as), max(as))+
  facet_grid(k0~eps,labeller = labeller) + labs(fill=TeX('\\xi')) + xlab(TeX('\\alpha')) + ylab(TeX('\\beta')) + theme_minimal()

```

The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.

For a given network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$ the likelihood is given by:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}

where $B(y,z)$ is the the beta function and $l\ge0$ variable that allows truncating the data such that the minimum degree is $l$.

This likelihood allows for the parameters to be inferred using either a frequentist or a Bayesian approach.

# Simulation Study {#sec-sim}

This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.

The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongisde the priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

to obtain a posterior distribution that can then be used in an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm to obtain posterior samples. The results of this inference are shown in @fig-rec1 and @fig-rec2. For these simulated networks $l=0$.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
#| label: fig-rec1

surfit_list = list()
pars = readRDS('results/recovery_pars.rds')
recovery_list = readRDS('results/recovery_dat.rds')
selected = c(1,27,14,35)
# selected = 1:36
for(k in 1:length(selected)){
  j=selected[k]
  x = recovery_list[[j]]$mcmc$dat[,1]
  ls = c()

  y_975 = recovery_list[[j]]$mcmc$surv$CI[2,][order(x)]
  y_025 = recovery_list[[j]]$mcmc$surv$CI[1,][order(x)]
  y_50 = recovery_list[[j]]$mcmc$surv$est[order(x)]
  x = sort(x)
  surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
    geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
    scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
    ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
                       ', $\\epsilon = $',pars$eps[j],
                       ', $\\beta = $',pars$b[j])))
}

fig = ggpubr::ggarrange(plotlist = surfit_list,
          nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
          label.x='Degree', label.y = 'Survival')

ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')

```

![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2}

@fig-rec1 and @fig-rec2 demonstrate that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network. This indicates that the method may also be able to be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.

# Application to Real Data {#sec-real}

Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @Lee24.

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-real1

plt1 = readRDS('results/comp_plot.rds')

plot(plt1)

```

@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model. In most cases, the GPA model does not necessarily provide an improvement in fit when compared to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme. @fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model. Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution and where it doesn't it appears to either be because it is fitting better at the tail than the Zipf-IGP model or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA. This again shows the effects that small degrees have on this model, which is somewhat expected as the theory used for this model is for trees and none of these real networks (nor many real networks) are.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-shapes

plt2 = readRDS('results/shape_plot.rds')

plot(plt2)

```

@fig-pa shows the estimated preference function $b$ alongside the 95% credible interval on a log-log plot. Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, for those that do the credible interval is much narrower as is the case for `sx-mathoverflow`. The most insightful conclusion we can draw from these plots come from the shape of the preference function. There appears to be two distinct shapes of preference function. The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`. The second distinct shape appears provide some clear PA behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimate for preference function (solid) with 95% credible interval (dashed) on log-log scale.
#| label: fig-pa

plt3 = readRDS('results/PA_plot.rds')
plot(plt3)

```

![: Posterior estimates of paramters for real data.](images/pars_plot.png)

# Conclusion and Discussion

In this paper we introduced a flexible class of preference function that when used under the GPA scheme (in the tree setting) is guaranteed to generate a network with a heavy tailed degree distribution whilst remaining flexible in the body. Using simulations from networks using this class of preference function we showed that the model parameters are quite easily estimated from the degrees alone at a snapshot in time. Seeing this we applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way. Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it came with the added benefit of giving a posterior estimate for a preference function.

Obviously this method had its flaws in that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks. Hopefully, as the field progresses we will be able to apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\newpage

# Proofs and Derivations {#sec-proofs .appendix}

## Proof of @prp-omega

Taking the form of the GPA degree survival: $$
\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
$$ and substituting into the formula for $\Omega(F,n)$:

\begin{align*}
\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
\end{align*}

Clearly if $b(n)=c$ or $\lim_{n\rightarrow\infty}b(n)=c$ for some $c>0$ then $\Omega(F,n)=0$. Now consider a non-constant $b(n)$ and re-write $\Omega(F,n)$ as:

\begin{align*}
\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}\\&+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\&+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
\end{align*}

Then if $\lim_{n\rightarrow\infty}b(n)=\infty$ it follows that:

\begin{align*}
\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} \\&- \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
\end{align*}

## Derivation of @eq-rho

For a preference function of the form:

$$
b(k) = \begin{cases}
g(k),&k<k_0\\
g(k_0) + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\beta>0, k_0\in\mathbb N$ we have that

\begin{align*}
\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
\end{align*}

Now using the fact that:

$$
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
$$ and reindexing the product in the second sum

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
\end{align*}

In order to simplify the infinite sum, consider:

\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{B}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
\end{align*}

this infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
\end{align*}

\newpage

# Additional Tables and Figures {.appendix}

\newpage

# References {.unnumbered}

<!-- the pattern below controls the placement of the references -->

::: {#refs}
:::