### 2025-06-23 paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-basic,
]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
  pdfkeywords={networks, discrete extremes, power law, preferential
attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Tail Flexibility in the Degrees of Preferential Attachment
Networks]{Tail Flexibility in the Degrees of Preferential Attachment
Networks}

% author setup
\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
% affil setup
\affil[1]{\orgdiv{School of Mathematics, Statistics and
Physics}, \orgname{Newcastle University}}

% abstract

\abstract{Devising the underlying generating mechanism of a real-life
network is difficult as, more often than not, only its snapshots are
available, but not its full evolution. One candidate for the generating
mechanism is preferential attachment which, in its simplest form,
results in a degree distribution that follows the power law.
Consequently, the growth of real-life networks that roughly display such
power-law behaviour is commonly modelled by preferential attachment.
However, the validity of the power law has been challenged by the
presence of alternatives with comparable performance, as well as the
recent findings that the right tail of the degree distribution is often
lighter than implied by the body, whilst still being heavy. In this
paper, we study a modified version of the model with a flexible
preference function that allows super/sub-linear behaviour whilst also
guaranteeing that the limiting degree distribution has a heavy tail. We
relate the distributions tail index directly to the model parameters,
allowing direct inference of the parameters from the degree distribution
alone.}

% keywords
\keywords{networks,  discrete extremes,  power law,  preferential
attachment}

\begin{document}
\maketitle

\newpage

\section{Introduction}\label{introduction}

Networks appear in many fields such as sociology, politics,
epidemiology, and economics. Statistical methods have been used to study
networks that appear in these fields and beyond, from the use of
stochastic block models to detect communities within the French
political blogosphere \citep{Latouche11}, to using Exponential Random
Graph Models (ERGMs) to analyse the global trade network
\citep{Setayesh22}, and the use of mechanistic models in neuroscience to
investigate the patterns of connections in neural systems
\citep{Betzel17}.

One example of mechanistic models is the preferential attachment (PA)
model, where as a network grows and new vertices join, they connect to
those already present in the network with probabilities proportional to
some preference function of their in-degrees. The well-known
Barabási-Albert (BA) model is a special case of this when \(b(k) = k+1\)
\citep{Barabasi99}.

These mechanistic models are generally useful when simulating networks,
but not for fitting to real data as usually only snapshots of a network
are available and not the full evolution - without this conclusions
about the growth are difficult to make. However, one statistic that is
usually available for snapshots of networks is the degree distribution,
making it valuable to study the degrees in a network. Often, it is
appealing to model the degree distributions of real networks using the
power law \footnote{Power law: \(\bar F(k) = L(k)\times k^{-1/\xi}\),
  where \(\xi>0\) and \(L(k)\) is slowly varying in \(k\).}, as it is
seemingly observed in much of the data. Furthermore, as the BA model
results in a power law degree distribution \citep{Barabasi99}, that the
power law fits well offers the interpretation that the network may have
grown according to PA or similar simple rules. However, attempting to
explain the networks growth using preferential attachment when the
degrees seemingly follow a power law may not be the best course of
action as \citet{krapivsky01} show that a general preference function
does not necessarily lead to power law degrees. Specifically, showing
that a preference function of form \(b(k) = k^\alpha\) leads to light
tails when \(\alpha<1\), a degenerate network (a finite number of
vertices end up with all of the added edges) when \(\alpha>1\), and only
yields heavy tails when \(\alpha=1\). \citet{rudas07} derives the
limiting degree distribution in terms of the preference function
\(b(\cdot)\) and its parameters, we build upon this by finding a class
of preference functions that result in realistic tails i.e.~heavy but
not as heavy as implied by the power law. On the other hand, it is still
being argued whether or not many real data actually follow the power
law. \citet{Broido_2019} provides some evidence that real world networks
rarely have degrees that follow a power law, through the use of
statistical analysis of almost 1000 networks across a wide variety of
fields; \citet{Voitalov_2019}, on the other hand, disagrees and claims
that these kinds of networks are not nearly as rare and only appear so
as a result of an unrealistic expectation of power law without
deviations or noise.

Extreme value methods have been scarcely used when it comes to analysis
of degree distributions and whether they follow the power law. The
absence of these methods essentially downweights the deviation of
vertices with the largest degrees, which are usually the most
influential, from the power law. \citet{Voitalov_2019} proposes
estimating the power law exponent of a degree distribution using three
estimators based in extreme value theory, in contrast to the estimators
generally used in network science, the degree distributions are then
split into one of three categories (not power-law, hardly power-law or
power-law) depending on the values of the estimates obtained from each
of the estimators. \citet{Lee24} proposes a framework for modelling
entire degree distributions using a combination of two generalisations
of the power law with a model selection step that is capable of
suggesting whether or not the data is adequately modelled by a power
law, with one conclusion being that most data seems to have a tail
lighter than what is implied by the power law in body whilst still being
heavy tailed. One of the models used by \citet{Lee24} is the Zipf-IGP
model where up until a certain threshold the data is modelled using the
Zipf disribution (a discrete power law) and beyond it a discrete version
of the Generalised Pareto distribution is used, we will later compare
our model to this as they share some similarities.
\citet{wang2022random} considers a preferential attachment model with an
affine preference function and reciprocal edges, finding that the joint
limiting distribution of the in-degrees and out-degrees is multivariate
regularly varying and has the property of hidden regular variation.

Evidently, the study of degree distributions is a very active but the
vast majority of research in the field (whether using extremes or not)
are purely descriptive in the sense that no information about the
preference function is revealed even if preferential attachment was the
underlying mechanism. This paper aims to address this gap, asking if the
degree distribution of a network is assumed to come from the
preferential attachment model, can we directly infer the model
parameters?

The remainder of this paper is as follows: Section 2 gives a detailed
description of the PA model alongside the theoretical results for the
limiting degree distribution with a focus on its tail behaviour in terms
of the preference function \(b(\cdot)\). It is shown that a heavy tailed
degree distribution can only be the result of an asymptotically linear
preference function ending with a suggestion for a preference function
that can guarantee heavy tails while remaining flexible.
Section~\ref{sec-model} utilises the preference function proposed at the
end of Section 2, and illustrates how the tail index of the degree
distribution varies with the model parameters. Section~\ref{sec-sim}
consists of a simulation study where networks are simulated from the
proposed model for various parameter combinations, demonstrating that
the parameters can be recovered from fitting the model to only the
degree distribution. Section~\ref{sec-real} fits the model to some real
data and provides posterior estimates for the preference function.
Section~\ref{sec-conc} concludes the article.

\newpage

\section{Tail Behaviour of Preferential Attachment
Model}\label{sec-tail}

The network generative model that we will focus on in this paper is the
General Preferential Attachment model in \citet{rudas07} and is defined
as follows:

Starting at time \(t=0\) with an initial network of \(m\) vertices that
each have no edges, at times \(t=1,2,\ldots\) a new vertex is added to
the network bringing with it \(m\) directed edges (with the new vertex
as the source); the target for each of these edges are selected from the
vertices already in the network with weights proportional to some
preference function \(b(\cdot)\) of their degree, where
\(b: N \mapsto \mathbb R^+\setminus\{0\}\) is such that:

\begin{equation}\phantomsection\label{eq-condb2}{
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
}\end{equation}

Special cases of this model include the Barabási-Albert (BA) model when
\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
with \(\xi=1/2\), and the Uniform Attachment (UA) model where \(b(k)=c\)
leading to a degree distribution that is light tailed.

Given condition \ref{eq-condb2} the survival function of the limiting
degree distribution, called the limiting survival hereafter, can be
analytically derived in the case where \(m=1\), obtained by considering
a branching process that is equivalent to the growth of the network, as
in \citet{rudas07}. Theorem 1 from \citet{rudas07} states that for the
tree \(\Upsilon(t)\) at time \(t\):

\begin{equation}\phantomsection\label{eq-survlim}{
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\)
and \(\hat\rho\) is the Laplace transform of the density of the point
process associated with the pure birth process that corresponds to a
nodes individual growth, that is:

\[
\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t.
\]

The limiting survival can be viewed as the limit of the empirical
proportion of vertices with degree over a threshold \(k\in\mathbb N\),
that is:

\[
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
\] which can also be written using Equation~\ref{eq-survlim} as:

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
\] Therefore, the corresponding probability mass function of the degree
distribution \(f(k) = \bar F(k-1) - \bar F(k)\) is

\[
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
\]

We are interested in how the tail behaviour of the discrete limiting
degree distribution is affected by the preference function \(b\), which
can be determined by the results of \citet{shimura12}.

For a distribution \(F\) with survival function \(\bar F\) and some
\(k\in\mathbb Z^+\) let

\[
\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}.
\]

\citet{shimura12} states that if
\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
then \(F\) is heavy-tailed with \(\bar F(k) \sim k^{-\alpha}\). On the
other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the
distribution is light-tailed. This allows us to show the following:

\begin{proposition}[]\protect\hypertarget{prp-omega}{}\label{prp-omega}

If \(\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}\) and
\(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\), then \[
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
\]

\end{proposition}

See Appendix \ref{sec-proofs} for the details of the proof.

Proposition~\ref{prp-omega} aligns with the result from
\citet{krapivsky01} demonstrating that a sub-linear preference function
will lead to a light-tailed distribution, as
\(\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0\) if \(b(k)=k^\alpha\) where
\(\alpha < 1\). Proposition 2.1 also aligns with the fact that BA model
produces a heavy-tailed degree distribution with index \(\xi=0.5\) by
considering the preference function \(b(k) = k + \varepsilon\), as
\(\lim_{k\rightarrow\infty}b(k+1)-b(k)=1\) leaving the tail index to be
\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\). So,
in order for the degree distribution to be heavy-tailed we need that the
limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is positive.
To determine the class of functions that will result in heavy-tailed
limiting degree distributions, we use the following result:

\begin{theorem}[Stolz-Cesàro Theorem
\citep{cesaro}]\protect\hypertarget{thm-stolz}{}\label{thm-stolz}

Let \((a_k)_{k\ge1}\) and \((b_k)_{k\ge1}\) be two sequence of real
numbers. Assume \((a_k)_{k\ge1}\) is a strictly monotone and divergent
sequence and the following limit exists:

\[
\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
\] Then,

\[
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
\]

\end{theorem}

\begin{proposition}[]\protect\hypertarget{prp-omega2}{}\label{prp-omega2}

Consider a GPA model with preference function \(b(\cdot)\) satisfying
Proposition~\ref{prp-omega}. Then the limiting degree distribution has a
heavy tail with index \(c/\lambda^{*}\) if and only if
\(\lim_{k\rightarrow\infty}b(k)/k = c > 0\).

\emph{Proof}

From Proposition~\ref{prp-omega}, we have that: \[
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
\] Now, setting \(b_k = b(k)\) and \(a_k = k\):

\[
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
\] meaning by Theorem~\ref{thm-stolz}:

\[
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
\]

\end{proposition}

Considering Proposition~\ref{prp-omega2} we propose the following
preference function:

\begin{equation}\phantomsection\label{eq-pref}{
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0,\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
}\end{equation} for \(\alpha,\beta, \varepsilon>0\) and
\(k_0\in\mathbb N\).

This preference function, as per Proposition~\ref{prp-omega}, will
produce a degree distribution with tail index \(\beta/\lambda^*\)
guaranteeing a heavy tail. We study this preference function further in
the next section.

\newpage

\section{Preferential Attachment with Flexible Heavy
Tail}\label{sec-model}

In the previous section, we found that using a preference function with
linearity in the limit allows for the inclusion of sub/super-linear
behaviour below the threshold, while simultaneously guaranteeing a heavy
tail. In this section, we study how the limiting degree distribution
varies with the preference function according to Equation~\ref{eq-pref}.
Its survival is

\begin{equation}\phantomsection\label{eq-polysurv}{
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
\end{cases}
}\end{equation}

with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where

\begin{equation}\phantomsection\label{eq-rho}{
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
}\end{equation}

which must be solved numerically for most parameter choices. Also, note
that \(\lambda>\beta\).

Some examples of the limiting degree distribution for various parameter
combinations are shown below on log-log scale in
Figure~\ref{fig-polylinsurv}:

\begin{verbatim}
     a   b  eps k0 x   lambda        surv
10 0.5 1.5 0.01 20 0 1.500094 0.006622105
\end{verbatim}

\begin{verbatim}
[1] 0.9999374
\end{verbatim}

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}

}

\caption{\label{fig-polylinsurv}Theoretical survival distributions of
the limiting degree distributions, according to various combinations of
\((\alpha, \beta, \varepsilon)\) and \(k_0=20\) of the proposed
preferential attachment model.}

\end{figure}%

Figure~\ref{fig-polylinsurv} demonstrates that this model can capture a
wide variety of shapes for the survival function, including a large
range of possible tail indices ranging from 0.035
(\(\alpha=1.5, \beta=0.1, \varepsilon=1\)) to 0.999
(\(\alpha=0.5, \beta=1.5, \varepsilon=0.01\)).

The survival function (\ref{eq-polysurv}) can be connected to the
discrete version of the generalised Pareto distribution (GP), called the
Integer GP (IGP) \citep{Rohrbeck_2018} with conditional survival: \[
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\), denoted as
\(X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)\) where \(\xi\) is the shape
parameter controlling the tail index

By Equation~\ref{eq-polysurv} and using Stirling's approximation:

\begin{align*}
\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
\end{align*}

Therefore, \[
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
\] meaning that for \(k\ge k_0\) the limiting degree distribution (for
large \(k_0^\alpha\)) is approximated by
\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).

To assess how close of an approximation this is, the theoretical
conditional survivals are shown in Figure~\ref{fig-approx_surv} in
colour and their IGP approximations are shown in grey. The approximation
seems to hold up fairly well even for large degrees and more so when
\(\alpha\) is larger.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}

}

\caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
alongside their IGP approximations (coloured).}

\end{figure}%

In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\) ensures that
the shape parameter of the IGP is positive and thus the distribution is
heavy-tailed. Additionally the value of the shape parameter \(\xi\) is
shown in Figure~\ref{fig-polyheat} for various parameter choices. The
darker and lighter regions on the heat maps correspond to a heavier and
a lighter tail, respectively, the red dashed line shows combinations of
\(\alpha\) and \(\beta\) that produce a limiting degree distribution
with the same tail index as the Barabási-Albert model, that is,
\(\xi=0.5\).

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}

}

\caption{\label{fig-polyheat}Heat maps of \(\xi\) for various
combinations of the parameters of the proposed model.}

\end{figure}%

\newpage

To perform inference of the model parameters, we consider a network with
degree count vector \(\pmb n = (n_0, n_1, \ldots, n_M)\), where \(M\) is
the maximum degree. We can then write the likelihood as:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}\label{eq-lh}

where \(B(y,z)\) is the the beta function, and \(l\ge0\) is a variable
that allows truncating the data such that the minimum degree is \(l\).
This will allow the model to be fitted whilst ignoring the influence of
the lower degrees (those less than \(l\)) as the model does not capture
the behaviour well at the lower degrees since \citet{rudas07} only
provides results for the case of a preferential attachment tree.

\section{Applications}\label{applications}

\subsection{Simulated Data}\label{sec-sim}

This first subsection aims to show that the parameters of the model in
Section~\ref{sec-model} can be recovered from simulating a network from
the model, and fitting it to the observed degree distribution, using the
likelihood in \eqref{eq-lh}.

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts and using the likelihood from the previous section alongisde the
priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

where Ga(a, b) is the Gamma distribution with shape a and rate b, and
U(a, b) is uniform distribution with lower and upper bounds a and b, to
obtain a posterior distribution, up to the proportionality constant.
Posterior samples can then be obtained by an adaptive
Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm. For these
simulated networks \(l=0\).

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}

}

\caption{\label{fig-rec1}Posterior estimates of survival function for
data simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/mcmc_plot.jpg}

}

\caption{\label{fig-rec2}: Posterior estimates of paramters for data
simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} shows the estimations
demonstrates the usefulness of the model, as we can recover the model
parameters fairly well from only the final degree distribution of the
simulated network. This indicates that the method may also be applied to
the degree distributions of real networks, estimating the model
parameters assuming they evolved according to the GPA scheme.

\subsection{Real Data}\label{sec-real}

In this section, we fit the proposed model to the degree distributions
of various real networks and learn about the mechanics of their growth.
While we also compare the fit to that of the mixture distribution by
\citet{Lee24} we note that the proposed method has the additional
benefit of learning directly about the growth of a network from the
inference results. The data consists of 12 networks sourced from
\href{konect.cc}{KONECT} and the
\href{https://networkrepository.com}{Network Data Repository}\citep{nr}:

\begin{itemize}
\tightlist
\item
  \texttt{as-caida20071105}: network of autonomous systems of the
  Internet connected with each other from the CAIDA project
\item
  \texttt{dimacs10-astro-ph} : co-authorship network from the
  ``astrophysics'' section (astro-ph) of arXiv
\item
  \texttt{ego-twitter}: network of twitter followers
\item
  \texttt{facebook-wosn-wall}: subset of network of Facebook wall pasts
\item
  \texttt{maayan-faa}: USA FAA (Federal Aviation Administration)
  preferred routes as recommended by the NFDC (National Flight Data
  Center)
\item
  \texttt{maayan-Stelzl}: network representing interacting pairs of
  proteins in Humans
\item
  \texttt{moreno-blogs-blogs}: network of URLs found on the first pages
  of individual blogs
\item
  \texttt{opsahl−openflights}: network containing flights between
  airports of the world.
\item
  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
\item
  \texttt{reactome}: network of protein--protein interactions in Humans
\item
  \texttt{sx-mathoverflow}: interactions from the StackExchange site
  \href{https://mathoverflow.net/}{MathOverflow}
\item
  \texttt{topology}: network of connections between autonomous systems
  of the Internet
\end{itemize}

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}

}

\caption{\label{fig-real1}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red).}

\end{figure}%

Figure~\ref{fig-real1} displays the posterior estimates of the survival
function for various data sets, obtained from fitting the GPA model and
the Zipf-IGP mixture model from \citet{Lee24}. In most cases, the GPA
model does not necessarily provide an improvement in fit when compared
to the Zipf-IGP model but where the GPA model fits well we gain
additional information about the preference function assuming that the
network evolved according the the GPA scheme.

Figure~\ref{fig-shapes} shows the posterior of the shape parameter
\(\xi\) obtained from the Zipf-IGP model alongside the posterior of the
equivalent shape parameter \(\beta/\lambda^*\) obtained from fitting the
GPA model. Generally, the GPA model performs similarly to the Zipf-IGP
when estimating the tail behaviour of the degree distribution. In the
cases of substantial discrepancies, it is either because the GPA model
fits the tail better than the Zipf-IGP model does, or because of the
threshold being estimated as too low forcing almost all of the data to
be modelled by the linear part of the GPA. This again shows the effects
that small degrees have on this model, which is somewhat expected as the
theory used for this model is for trees and none of these real networks
(nor many real networks) are.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}

}

\caption{\label{fig-shapes}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red).}

\end{figure}%

Figure~\ref{fig-pa} shows the estimated preference function \(b(k)\)
alongside the 95\% credible interval on a log-log plot. Although the
credible interval becomes very large for the largest degrees, this is
expected as not all of these networks had data in that region, and for
those that do the credible interval is much narrower, as is the case for
\texttt{sx-mathoverflow}. Looking at the shape of the preference
function, there appears to be two distinct shapes of preference
function. The first appears mostly flat (similar to uniform attachment)
for the smallest degrees and then after a threshold preferential
attachment kicks in, some with this shape are \texttt{pajek-erdos} and
\texttt{sx-mathoverflow}. The second distinct shape appears to provide
some clear preferential attachment behaviour that then slows down after
a certain point, examples of this are seen in the two infrastructure
networks \texttt{opsahl-openflights} and \texttt{topology}. This slowing
down could be viewed as a kind of diminishing returns on the degree of a
vertex i.e.~as a vertex gets larger gaining more connections has less of
an effect than it did before some threshold \(k_0\).

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}

}

\caption{\label{fig-pa}Posterior estimate for preference function
(solid) with 95\% credible interval (dashed) on log-log scale.}

\end{figure}%

\newpage

\section{Conclusion and Discussion}\label{sec-conc}

In this paper we introduced a class of preference functions that, under
the GPA scheme, generate a network with a flexible yet heavy-tailed
degree distribution. From the simulation study we showed that the
parameters can be recovered from fitting the model to the degrees alone.
We also applied this method to the degree distributions of real
networks, estimating their model parameters assuming they evolved in the
same way. Not only did this yield fairly good fits for the degree
distribution, similar to that of the Zipf-IGP, it also came with the
added benefit of giving a posterior estimate for a preference function.

One limitation of this method is that the lowest degrees needed to be
truncated as they had a very large effect on the fit of the model as a
result of using theory developed for trees and applying it to general
networks. Future work could apply theory developed for general networks
using a similar method to this, allowing us to compare the results here
something that is more accurate. This could include fixing the
out-degree of new nodes at a constant greater than one, or allowing the
out-degree of new nodes to vary.

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\newpage

\section{Additional Plots}\label{additional-plots}

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.png}

}

\caption{: Posterior estimates of parameters for real data.}

\end{figure}%

\section{Proofs and Derivations}\label{sec-proofs}

\subsection{\texorpdfstring{Proof of
Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}

Taking the form of the GPA degree survival: \[
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda+b(i)}
\] and substituting into the formula for \(\Omega(F,n)\):

\begin{align*}
\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda+b(k+1)}{b(k+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}.
\end{align*}

Clearly if \(b(k)=c\) or \(\lim_{k\rightarrow\infty}b(k)=c\) for some
\(c>0\) then \(\Omega(F,k)=0\). Now consider a non-constant \(b(k)\) and
re-write \(\Omega(F,k)\) as:

\begin{align*}
\Omega(F,k) &= \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}+\frac{b(k+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}\\&+\frac{b(k+1)}{\lambda}  -\frac{b(k+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\&+\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}.
\end{align*}

Then if \(\lim_{k\rightarrow\infty}b(k)=\infty\) it follows that:

\begin{align*}
\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} \\&- \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\
&\qquad+\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad \square
\end{align*}

\subsection{\texorpdfstring{Derivation of
Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{derivation-of-eq-rho}

For a preference function of the form:

\[
b(k) = \begin{cases}
g(k),&k<k_0,\\
g(k_0) + \beta(k-k_0), &k\ge k_0,
\end{cases}
\] for \(\beta>0, k_0\in\mathbb N\) we have that

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)}\\ &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}.
\end{align*}

Now using the fact that:

\[
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
\] and reindexing the product in the second sum,

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}.
\end{align*}

In order to simplify the infinite sum, consider:

\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{B}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}.
\end{align*}

This infinite sum does not converge when \(x\le1\) as each term is
\(O(n^{-x})\). We can now use this in \(\hat\rho(\lambda)\):

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}.\qquad \square
\end{align*}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{refs.bib}

\end{document}