### 2025-08-28 paper.tex
Making edits

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-basic,
]{sn-jnl}

\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
  pdfkeywords={networks, discrete extremes, power law, preferential
attachment},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Learning growth mechanisms of tail realistic preferential
attachment models from network degree distributions]{Learning growth
mechanisms of tail realistic preferential attachment models from network
degree distributions}

% author setup
\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
% affil setup
\affil[1]{\orgdiv{School of Mathematics, Statistics and
Physics}, \orgname{Newcastle University}}

% abstract

\abstract{Devising the underlying generating mechanism of a real-life
network is difficult as, more often than not, only its snapshots are
available, but not its full evolution. One candidate for the generating
mechanism is preferential attachment which, in its simplest form,
results in a degree distribution that follows the power law.
Consequently, the growth of real-life networks that roughly display such
power-law behaviour is commonly modelled by preferential attachment.
However, the validity of the power law has been challenged by the
presence of alternatives with comparable performance, as well as the
recent findings that the right tail of the degree distribution is often
lighter than implied by the body, whilst still being heavy. In this
paper, we study a modified version of the model with a flexible
preference function that allows super/sub-linear behaviour whilst also
guaranteeing that the limiting degree distribution has a heavy tail. We
relate the distributions tail index directly to the model parameters,
allowing direct inference of the parameters from the degree distribution
alone.}

% keywords
\keywords{networks,  discrete extremes,  power law,  preferential
attachment}

\begin{document}
\maketitle

\newpage

\section{Introduction}\label{introduction}

Networks have become very powerful tools for representing and analysing
complex systems, with uses in a large array of fields. In network
science and statistics, they have been studies by various families of
models, from stochastic block models for detecting communities online
\citep{Latouche11}, to exponential random graph models (ERGMs) for
analysing the global trade network \citep{Setayesh22}, and mechanistic
models for investigating patterns in neural systems \citep{Betzel17}.

Amid the recent rise of interest in networks, there has been a debate on
whether most real networks are scale-free. Claiming a real network is
scale-free is equivalent to saying that its degree distribution follows
a power law, that is the fraction of nodes with degree \(k\) is
proportional to \(k^{-\alpha}\), and therefore has a regularly varying
tail with tail index \(\alpha\). On the side against the claim is
\citet{Broido_2019who} compared the fits of a power law model against
that of several non-scale-free models to nearly a thousand networks,
only to find strong evidence for scale-freeness in four percent and weak
evidence in over half of the networks, thus claiming that scale-free
networks do not make up a majority in real networks. They compare the
fits of a power law model against that of several non-scale-free models
only finding strong evidence for scale-freeness in four percent and weak
evidence in fifty-two percent of the networks. On the other side of this
debate is \citet{Voitalov_2019} who disagrees and claims that these
networks are not nearly as rare and only appear to so as a result of an
unrealistic expectation of a power law without deviations or noise.
Additional evidence of these deviations from a power law is shown by
\citet{Lee24} who demonstrate that a lot of networks are partially
scale-free, in that the body of the degree distribution is often
modelled well by a power law, while the tail is lighter than what is
implied by the body, albeit still regularly varying..

Nevertheless, most studies into the appropriateness of a power law for
the degrees of real networks, the aforementioned references included,
have been largely descriptive in the sense that no information about the
growth of the networks is revealed.

The popularity of using the power law for network degrees can be traced
back to the preferential attachment (PA) model popularised by
\citet{Barabasi99}. In the general model, as new nodes join the network,
an existing node with degree k gains edges at a rate proportional to
\(b(k)\), where b is a non-negative preference function.
\citet{Barabasi99} showed that, when \(b(k) = k + 1\), in the limit the
resulting degree distribution is regularly varying with index 3.
Subsequently, if a real network is shown to be scale-free, one can
loosely justify preferential attachment as the underlying mechanism of
its growth.

The model from \citet{Barabasi99} provided the foundations for various
generalisations.

\citet{krapivsky01} considered \(b(k) = (k+1)^\alpha\), and showed that
the degree distribution is not regularly varying (and therefore not
following the power law) when \(0<\alpha<1\), and when \(\alpha>1\) a
finite number of nodes end up with all edges after a certain point
resulting in a degenerate degree distribution. \citet{wang2022random}
returns to a linear preference function of the form
\(b(k) = k+\varepsilon\) but adds the possibility for reciprocal edges
to be sent, resulting in the joint distribution of in-degree and
out-degree being multivariate regularly varying and having the property
of hidden regular variation. \citet{rudas07} follows in the footsteps of
\citet{krapivsky01} , by considering a preferential attachment tree and
using theory from continuous branching processes, derives a limiting
degree distribution in terms of the preference function \(b(\cdot)\).
Nevertheless, research in this area tends to only focus on the
theoretical asymptotic results of network growth models with little
analysis of real networks.

This paper aims to address the gap between the applied and theoretical
works, by asking if a network is assumed to come from a preferential
attachment model, can we use the degree distribution alone to directly
infer the parameters of the the preference function and learn about the
growth mechanisms? Moreover, proper consideration is given to the tail
of the degree distribution, because otherwise the effects of the largest
degrees, which correspond to the most influential nodes, deviating from
the power law will be discounted.

It is important to think about how we intend to consider the tail of the
degree distributions, because if not done correctly we may end up
essentially discounting the effects of the largest degrees (often the
most influential nodes) deviating from a power law.

For the above reason, we will use methods from discrete extreme value
theory, in particular those by \citet{shimura12}, who provided
theoretical guarantees for a discrete distribution to be regularly
varying or not. Using these results, we demonstrate how the tail of the
degree distribution is affected by \(b(\cdot)\), and subsequently
propose a class of preference functions that is tail realistic for real
networks. These analytical results enables the likelihood of the degree
distribution to be expressed in terms of the parameters of \(b(\cdot)\),
which in turn allows the underlying mechanism of the network, assumed to
grow according to preferential attachment, to be inferred directly.

The remainder of the paper is as follows: Section~\ref{sec-tail} gives a
detailed description of the preferential attachment model alongside the
theoretical results for the survival of the limiting degrees, with a
focus on the tail behaviour in terms of the preference function
\(b(\cdot)\). A class of asymptotically linear preference functions will
be introduced and shown to guarantee regular variation in the degrees
while remaining flexible up until a threshold. Section~\ref{sec-model}
utilises the proposed preference function and illustrates numerically
how the tail index of the degree distribution varies with the model
parameters. Section~\ref{sec-sim} consists of a simulation study where
networks are simulated from the proposed model for various parameter
combinations, demonstrating that the parameters can be recovered from
fitting the model to only the degree distribution.
Section~\ref{sec-real} fits the model to some real data and provides
posterior estimates for the preference function. Section~\ref{sec-conc}
concludes the article.

\newpage

\section{Tail Behaviour of Preferential Attachment
Model}\label{sec-tail}

The model that we will focus on in this paper is the General
Preferential Attachment model in \citet{rudas07} and is defined as
follows:

Starting at time \(t=0\) with an initial network of \(m\) vertices that
each have no edges, at times \(t=1,2,\ldots\) a new vertex is added to
the network bringing with it \(m\) directed edges from the new vertex;
the target for each of these edges are selected from the vertices
already in the network with weights proportional to some non-decreasing
preference function \(b(\cdot)\) of their degree, where
\(b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}\) is such that:

\begin{equation}\phantomsection\label{eq-condb2}{
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
}\end{equation}

Special cases of this model include the Barabási-Albert (BA) model when
\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
with tail index 2, and the Uniform Attachment (UA) model where
\(b(k)=c\) leading to a degree distribution that is not regularly
varying.

The survival function of the limiting degree distribution, called the
limiting survival hereafter, under condition \ref{eq-condb2} can be
analytically derived in the case where \(m=1\), this is presented below.

Consider a continuous time branching process \(\zeta(t)\) driven by a
Markovian pure birth process, with \(\zeta(0)=0\) and birth rates
depending on a non-negative function \(b(\cdot)\):

\[
\Pr(\zeta(t+dt)=k+1|\zeta(t)=k) = b(k)dt + o(dt).
\] Now let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
follows: \(\Upsilon(t)=\{\emptyset\}\) and \(\Upsilon(t)=G\) where each
existing node \(x\) in \(\Upsilon(t)\) gives birth to a child with rate
\(b(\mathrm{deg}(x, \Upsilon(t))\) independently of the other nodes
where \(\mathrm{deg}(x, \Upsilon(t))\) is the degree of node \(x\) in
the tree \(\Upsilon(t)\) at time \(t\).

Theorem 1 from \citet{rudas07} states that for the tree \(\Upsilon(t)\)
at time \(t\) and a characteristic function of the tree
\(\varphi(\cdot)\) :

\begin{equation}\phantomsection\label{eq-survlim}{
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\)
and \(\hat\rho\) is the Laplace transform of the density of the point
process associated with the pure birth process that corresponds to the
growth of an individual node, that is
\(\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t\)

The limiting survival can be viewed as the limit of the empirical
proportion of vertices with degree over a threshold \(k\in\mathbb N\),
that is:

\[
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
\] which can also be written using Equation~\ref{eq-survlim} as:

\begin{equation}\phantomsection\label{eq-surv-origin}{
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
}\end{equation}

Therefore, the corresponding probability mass function of the degree
distribution \(f(k) = \bar F(k-1) - \bar F(k)\) is

\begin{equation}\phantomsection\label{eq-pmf-origin}{
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
}\end{equation}

We are interested in how the tail behaviour of the discrete limiting
degree distribution is affected by the preference function \(b\).

For a distribution \(F\) with survival function \(\bar F\) and
\(k\in\mathbb Z^+\) let

\[
\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}.
\]

This allows us to show the following:

\begin{proposition}[]\protect\hypertarget{prp-omega}{}\label{prp-omega}

If \(\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}\) and
\(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\), then \[
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
\]

\end{proposition}

See Appendix \ref{sec-proofs} for the details of the proof.

\citet{shimura12} states that if
\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\). On
the other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then we
will refer to the distribution as light-tailed.

Proposition~\ref{prp-omega} aligns with the result from
\citet{krapivsky01} demonstrating that a sub-linear preference function
will lead to a light-tailed distribution, as
\(\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0\) if \(b(k)=k^\alpha\) where
\(\alpha < 1\). Proposition 2.1 also aligns with the fact that BA model
produces a regularly varying degree distribution, with tail index 2, by
considering the preference function \(b(k) = k + \varepsilon\), as
\(\lim_{k\rightarrow\infty}b(k+1)-b(k)=1\) leaving the tail index to be
\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\). So,
in order for the degree distribution to be regularly varying we need
that the limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is
positive. The following proposition determines the class of functions
that will result in regular varying degree distributions.

\begin{proposition}[]\protect\hypertarget{prp-omega2}{}\label{prp-omega2}

Consider a GPA model with preference function \(b(\cdot)\) satisfying
Proposition~\ref{prp-omega}. Then the limiting degree distribution is
regularly varying with tail index \(\lambda^{*}/c\) if and only if
\(\lim_{k\rightarrow\infty}b(k)/k = c > 0\).

\emph{Proof}

From Proposition~\ref{prp-omega}, we have that: \[
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
\] Now, setting \(b_k = b(k)\) and \(a_k = k\):

\[
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
\] Using the Stolz-Cesaro Theorem
\href{Appendix\%20-@sec-sup}{@thm-stolz},

\[
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
\]

\end{proposition}

Using this result we can understand how the preference function is
directly connected with the tail behaviour of the degree distribution.
Specifically, regular variation is achieved if and only if \(b(k)\) is
asymptotically linear with \(k\). We use this result to create a
preference function that guarantees regular variation in the tail of the
degree distribution, aligning with analysis of real networks, whilst
allowing for the tail to deviate from the shape implied by the body.
This gives the model the capability to produce realistic behaviour in
the degrees like what was in \citet{Lee24} by using a piecewise function
inspired by the observed deviation from the power law after a certain
threshold:

\begin{equation}\phantomsection\label{eq-pref}{
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0,\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
}\end{equation} for \(\alpha,\beta, \varepsilon>0\) and
\(k_0\in\mathbb N\).

Using Proposition~\ref{prp-omega}, we can show that
\(\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta\), meaning the
degree distribution obtained using this preference funnction is
regularly varying with tail index \(\lambda^*/\beta\).

\newpage

\section{A Preferential Attachment Model with Flexible Regular
Variation}\label{sec-model}

In the previous section, we found that using an asymptotically linear
preference function allows for the inclusion of sub/super-linear
behaviour below the threshold, while simultaneously guaranteeing regular
varation of the degrees. In this section, we demonstrate the flexibility
of the preference function in Equation~\ref{eq-pref}, with regard to the
tail behaviour of the limiting degree distribution. Using
Equation~\ref{eq-surv-origin}, the limiting survival is

\begin{equation}\phantomsection\label{eq-polysurv}{
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
\end{cases}
}\end{equation}

with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where

\begin{equation}\phantomsection\label{eq-rho}{
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
}\end{equation}

which has to be solved numerically for most parameter choices. Also,
note that \(\lambda>\beta\).

For some parameter combinations, the limiting survival \(\bar F(k)\) is
shown on log-log scale in Figure~\ref{fig-polylinsurv}:

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}

}

\caption{\label{fig-polylinsurv}The limiting survival, according to
various combinations of \((\alpha, \beta, \varepsilon)\) and \(k_0=20\)
of the proposed preferential attachment model.}

\end{figure}%

Figure~\ref{fig-polylinsurv} demonstrates that this model can capture a
range of tail behaviour, including a large range of possible tail
indices ranging from 0.035 (\(\alpha=1.5, \beta=0.1, \varepsilon=1\)) to
0.999 (\(\alpha=0.5, \beta=1.5, \varepsilon=0.01\)).

The analytic form of the survival function in (\ref{eq-polysurv}),
offers a natural connection to the discrete version of the generalised
Pareto distribution (GP), providing a link to a well established
component of discrete extremes in the literature. Specifically,
connecting to the Integer GP (IGP) seen in \citet{Rohrbeck_2018} with
conditional survival: \[
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\), denoted as
\(X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)\) where \(\xi\) is the shape
parameter controlling the tail index.

By Equation~\ref{eq-polysurv} and using Stirling's approximation:

\begin{align*}
\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}.
\end{align*}

Therefore, \[{#eq-igp-est}
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
\] meaning that for \(k\ge k_0\) the limiting degree distribution (for
large \(k_0^\alpha\)) is approximated by the
\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).

To assess how close of an approximation this is, the theoretical
conditional survival Equation~\ref{eq-polysurv} are shown in
Figure~\ref{fig-approx_surv} in colour and their IGP approximations
\textbf{?@eq-igp-est} are shown in grey. The approximation seems to hold
up fairly well even for large degrees and more so when \(\alpha\) is
larger.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}

}

\caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
alongside their IGP approximations (coloured).}

\end{figure}%

In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\) ensures that
the shape parameter of the IGP is positive and thus the distribution is
regularly varying. Additionally the value of the shape parameter \(\xi\)
is shown in Figure~\ref{fig-polyheat} for various parameter choices. The
darker and lighter regions on the heat maps correspond to a heavier and
a lighter tail, respectively, the red dashed line shows combinations of
\(\alpha\) and \(\beta\) that produce a limiting degree distribution
with the same tail index as the Barabási-Albert model.

Through the connection to the IGP, fitting this model is almost
equivalent to fitting the IGP to the degrees and estimating its
parameters. However, instead of only describing the shape of the degree
distribution, we would also gain estimates for the shape of the
preference function, thus gaining a direct understanding the mechanisms
underlying the growth of the network.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}

}

\caption{\label{fig-polyheat}Heat maps of \(\xi\) for various
combinations of the parameters of the proposed model.}

\end{figure}%

\newpage

To perform inference of the model parameters, we consider a network with
degree count vector \(\pmb n = (n_0, n_1, \ldots, n_M)\), where \(M\) is
the maximum degree. Using Equation~\ref{eq-pmf-origin}, the likelihood
is:

\begin{align*}
L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}\label{eq-lh}

where \(\text{B}(\cdot,\cdot)\) is the the beta function, and \(l\ge0\)
is a quantity that allows truncating the data such that the minimum
degree is \(l\). This will allow the model to be fitted whilst ignoring
the influence of the lower degrees (those less than \(l\)) as the model
does not capture the behaviour well at the lower degrees since
\citet{rudas07} only provides results for the case of a preferential
attachment tree.

\section{Applications}\label{applications}

\subsection{Simulated Data}\label{sec-sim}

This subsection aims to show that the parameters of the model (and
therefore the preference function) in Section~\ref{sec-model} can be
recovered from simulating a network from the model, and fitting it to
the observed degree distribution, using the likelihood in \eqref{eq-lh}.

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts and using the likelihood from the previous section alongisde the
priors:

\begin{align*}
\alpha&\sim \text{Gamma}(1,0.01),\\
\beta &\sim  \text{Gamma}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Gamma(1,0.01)},
\end{align*}

where Gamma(\(a\), \(b\)) is the Gamma distribution with shape \(a\) and
rate \(b\), and U(\(a\), \(b\)) is uniform distribution with lower and
upper bounds a and b, to obtain a posterior distribution, up to the
proportionality constant. Posterior samples can then be obtained by an
adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
For these simulated networks \(l=0\).

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}

}

\caption{\label{fig-rec1}Posterior estimates of survival function for
data simulated from the proposed model with various combinations of
\((\alpha,\beta,\varepsilon)\) and \(k_0=20\).}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{images/mcmc_plot.jpg}

}

\caption{\label{fig-rec2}: Posterior estimates of paramters for data
simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

Figures \ref{fig-rec1} and \ref{fig-rec2} demonstrate the usefulness of
the model, as we can recover the model parameters well from only the
final degree distribution of the simulated network. This indicates that
the method may also be applied to real networks, with the assumption
that they evolved according to the GPA scheme.

\subsection{Real Data}\label{sec-real}

In this section, we fit the proposed model to the degree distributions
of various real networks and learn about the mechanics of their growth.
While we also compare the fit to that of the mixture distribution by
\citet{Lee24} we note that the proposed method has the additional
benefit of learning directly about the growth of a network from the
inference results. The data consists of 12 networks sourced from
\href{konect.cc}{KONECT} and the
\href{https://networkrepository.com}{Network Data Repository}\citep{nr}:

\begin{itemize}
\tightlist
\item
  \texttt{as-caida20071105}: network of autonomous systems of the
  Internet connected with each other from the CAIDA project
\item
  \texttt{dimacs10-astro-ph} : co-authorship network from the
  ``astrophysics'' section (astro-ph) of arXiv
\item
  \texttt{ego-twitter}: network of twitter followers
\item
  \texttt{facebook-wosn-wall}: subset of network of Facebook wall posts
\item
  \texttt{maayan-faa}: USA FAA (Federal Aviation Administration)
  preferred routes as recommended by the NFDC (National Flight Data
  Center)
\item
  \texttt{maayan-Stelzl}: network representing interacting pairs of
  proteins in humans
\item
  \texttt{moreno-blogs-blogs}: network of URLs found on the first pages
  of individual blogs
\item
  \texttt{opsahl−openflights}: network containing flights between
  airports of the world.
\item
  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
\item
  \texttt{reactome}: network of protein--protein interactions in humans
\item
  \texttt{sx-mathoverflow}: interactions from the StackExchange site
  \href{https://mathoverflow.net/}{MathOverflow}
\item
  \texttt{topology}: network of connections between autonomous systems
  of the Internet
\end{itemize}

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}

}

\caption{\label{fig-real1}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red).}

\end{figure}%

Figure~\ref{fig-real1} displays the posterior estimates of the survival
function for various data sets, obtained from fitting the GPA model and
the Zipf-IGP mixture model from \citet{Lee24}. In most cases, the GPA
model gives a similar fit to the Zipf-IGP model but where the GPA model
fits well we gain additional information about the preference function
assuming that the network evolved according the the GPA scheme.

Figure~\ref{fig-shapes} shows the posterior of the shape parameter
\(\xi\) obtained from the Zipf-IGP model alongside the posterior of the
equivalent shape parameter \(\beta/\lambda^*\) obtained from fitting the
GPA model. Generally, the GPA model performs similarly to the Zipf-IGP
when estimating the tail behaviour of the degree distribution. In the
cases of substantial discrepancies, it is either because the GPA model
fits the tail better than the Zipf-IGP model does, or because of the
threshold being estimated as too low forcing almost all of the data to
be modelled by the linear part of the GPA. This again shows the effects
that small degrees have on this model.

\begin{figure}

\centering{

\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}

}

\caption{\label{fig-shapes}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red).}

\end{figure}%

Figure~\ref{fig-pa} shows the estimated preference function \(b(k)\)
alongside the 95\% credible interval on a log-log scale. Although the
credible interval becomes very large for the largest degrees, this is
expected as not all of these networks had data in that region, and for
those that do the credible interval is much narrower, as is the case for
\texttt{sx-mathoverflow}. Looking at the shape of the preference
function, there appears to be two distinct shapes of preference
function. The first appears mostly flat (similar to uniform attachment)
for the smallest degrees and then after a threshold preferential
attachment kicks in, some with this shape are \texttt{pajek-erdos} and
\texttt{sx-mathoverflow}. The second distinct shape appears to provide
some clear preferential attachment behaviour that then slows down after
a certain point, examples of this are seen in the two infrastructure
networks \texttt{opsahl-openflights} and \texttt{topology}. This slowing
down could be viewed as a kind of diminishing returns on the degree of a
vertex i.e.~as a vertex gets larger gaining more connections has less of
an effect than it did before some threshold \(k_0\).

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}

}

\caption{\label{fig-pa}Posterior estimate for preference function
(solid) with 95\% credible interval (dashed) on log-log scale.}

\end{figure}%

\newpage

\section{Conclusion and Discussion}\label{sec-conc}

In this paper we introduced a class of preference functions that, under
the GPA scheme, generate a network with a flexible yet regularly varying
degree distribution. From the simulation study we showed that the
parameters can be recovered from fitting the model to the degrees alone.
We also applied this method to the degree distributions of real
networks, estimating their model parameters assuming they evolved in the
same way. Not only did this yield fairly good fits for the degree
distribution, similar to that of the Zipf-IGP, it also came with the
added benefit of giving a posterior estimate for a preference function.

One limitation of this method is that the lowest degrees needed to be
truncated as they had a very large effect on the fit of the model as a
result of using theory developed for trees and applying it to general
networks. Future work could apply theory developed for general networks
using a similar method to this, allowing us to compare the results here
something that is more accurate. This could include fixing the
out-degree of new nodes at a constant greater than one, or allowing the
out-degree of new nodes to vary.

Its worth noting the recent work by \citet{banerjee25}, that provides
results for more general networks beyond trees utilising the same
underlying branching process. They consider a model that grows in the
exact same way as the model that we have considered, but then they
collapse the nodes of the tree resulting in a more general network much
more like those in reality. However, there is no expression for the
probability mass function of the degrees and therefore no likelihood
that can be used for modeling in the same way that we have. In spite of
this, using Remark 2 from \citet{banerjee25} it can be shown that the
degree distribution is still heavy tailed although not necessarily
regularly varying.

Obtaining an expression for the degree distributions of more general
cases of preferential attachment models are, for the moment, seemingly
inaccessible and provide a real barrier to using more realistic models
in a way similar to what we have in this paper.

\setcounter{section}{0}
\renewcommand{\thesection}{\Alph{section}}
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

\newpage

\section{Supplementary Results}\label{sec-sup}

\begin{theorem}[Stolz-Cesàro Theorem
\citep{cesaro}]\protect\hypertarget{thm-stolz}{}\label{thm-stolz}

Let \((a_k)_{k\ge1}\) and \((b_k)_{k\ge1}\) be two sequence of real
numbers. Assume \((a_k)_{k\ge1}\) is a strictly monotone and divergent
sequence and the following limit exists:

\[
\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
\] Then,

\[
\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
\]

\end{theorem}

\section{Additional Plots}\label{additional-plots}

\begin{figure}[H]

{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.png}

}

\caption{: Posterior estimates of parameters for real data.}

\end{figure}%

\section{Proofs and Derivations}\label{sec-proofs}

\subsection{\texorpdfstring{Proof of
Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}

Taking the form of the GPA degree survival from
Equation~\ref{eq-surv-origin} : \[
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^*+b(i)}
\] and substituting into the formula for \(\Omega(F,n)\):

\begin{align*}
\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda+b(k+1)}{b(k+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}.
\end{align*}

Clearly if \(\lim_{k\rightarrow\infty}b(k)=c\) for some \(c>0\) then
\(\Omega(F,k)= 0 (= \lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}} [b(k+2)-b(k+1)])\).
Now consider a non-constant \(b(k)\) and re-write \(\Omega(F,k)\) as:

\begin{align*}
\Omega(F,k) &= \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}+\frac{b(k+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}\\&+\frac{b(k+1)}{\lambda}  -\frac{b(k+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\&+\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}.
\end{align*}

Then if \(\lim_{k\rightarrow\infty}b(k)=\infty\) it follows that:

\begin{align*}
\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} \\&- \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\
&\qquad+\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad \square
\end{align*}

\subsection{\texorpdfstring{Derivation of
Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{derivation-of-eq-rho}

For a preference function of the form:

\[
b(k) = \begin{cases}
g(k),&k<k_0,\\
g(k_0) + \beta(k-k_0), &k\ge k_0,
\end{cases}
\] for \(\beta>0, k_0\in\mathbb N\) we have that

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)}\\ &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}.
\end{align*}

Now using the fact that:

\[
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma\left(\frac{x}{y}+n+1\right)}{\Gamma\left(\frac{x}{y}\right)}
\] and reindexing the product in the second sum,

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}.
\end{align*}

In order to simplify the infinite sum, consider:

\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{B}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}.
\end{align*}

This infinite sum does not converge when \(x\le1\) as each term is
\(O(n^{-x})\). We can now use this in \(\hat\rho(\lambda)\):

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}.\qquad \square
\end{align*}

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{refs.bib}

\end{document}