### 2025-09-22 Files/archive/paper.qmd
Restructuring

---
title: Tail Flexibility in the Degrees of Preferential
 Attachment Networks
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-basic
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    number-sections: true
author:
  - name: Thomas William Boughen
    affiliations:
      - name: Newcastle University
        id: 1
        department: School of Mathematics, Statistics and Physics
  - name: Clement Lee
    affiliations:
      - ref: 1
  - name: Vianey Palacios Ramirez
    affiliations:
      - ref: 1
abstract: |
  Devising the underlying generating mechanism of a real-life network is difficult as, more often than not, only its snapshots are available, but not its full evolution. One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law. Consequently, the growth of real-life networks that roughly display such power-law behaviour is commonly modelled by preferential attachment. However, the validity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the right tail of the degree distribution is often lighter than implied by the body, whilst still being heavy. In this paper, we study a modified version of the model with a flexible preference function that allows super/sub-linear behaviour whilst also guaranteeing that the limiting degree distribution has a heavy tail. We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
keywords: [networks, extremes]
bibliography: refs.bib

---
```{r,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(paletteer)
library(gridExtra)
library(latex2exp)
library(ggpubr)
library(Rcpp)
library(RcppArmadillo)
```

# Introduction {#sec-intro}

Networks appear in many fields such as sociology, politics, epidemiology, and economics. Statistical methods have been and continue to be used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17]. One such mechanistic model is the general preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b$ of their degree. The well-known Barabási-Albert (BA) model is a special case of this when $b\sim k$ [BA99].

Mechanistic models, like the PA model, are in general only particularly useful when simulating networks and not fitting to the real data as usually only snapshots of real networks are available and not the entire evolution. As such, one of the most informative statistics when regarding a network is the degree distribution. When looking at the degree distributions of real networks it is attractive to many to use the power law to model the degree distribution, as the BA model generates degrees that follow a power law, and therefore claim that since the degree distribution is roughly power law then the underlying growth mechanism may be preferential attachment such as the BA model.  However, it has been shown that a general preference function will not always lead to a power law degree distribution [@krapivsky01] and additionally it has been debated for decades whether most real data actually follows a power law [@Voitalov_2019] or not[@Broido_2019].

Throughout this debate the use of methods from extremes has been absent until relatively recently with the papers such as [@Voitalov_2019] and [Lee]. Forgoing the use of extreme value methods essentially downplays how deviations from the power law in the largest degrees may affect the network given that they are often much more influential than the vertices with smaller degrees.

Noting this, many have now started to use these extreme value methods when analysing networks. [@wan19] showed that extreme value estimation methods prove more robust than existing parametric approaches especially when the data are corrupted or the model has been misspecified, demonstrating that extreme value methods can be a valuable tool when studying networks. [@wang2022random] investigates a PA model with a affine preference function (similar to the BA model) with the added behaviour of heterogeneous reciprocal edges, they show that the distribution of in-degrees and out-degrees produced by such a model is multivariate regularly varying and they also have hidden regular variation. [Lee] using methods from discrete extremes modelled the right tail of network degrees with a  discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:

$$
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
$$
for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.

The study of degree distributions, whether using extremes or not, usually does not reveal any information about the preference function even if PA was the underlying generating mechanism of the network. Here we address this gap in the literature, more specifically investigating if given the degree distribution of a network assumed to come from the PA model can we directly infer the model parameters? 
We build upon the foundation provided by [@rudas07] deriving the limiting degree distribution in terms of the preference function and suggest a class of preference functions that result in realistic tails that are heavy but are lighter than implied by the power law.

## Structure of this paper

@sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body. @sec-rec shows that the preference function parameters can be fairly well recovered from the degree distributions of networks simulated from the GPA model using various functions of the class introduced in @sec-gpa. Having shown that for simulated data the model parameters are recoverable from the degree distribution alone, @sec-real builds upon this and attempts to fit the model to degree distributions of real networks; modelling not only the degree distribution itself but also the estimates for the parameters of the preference function assuming the networks evolved according to the GPA scheme. Ultimately, @sec-conc discusses the main results of the paper, pitfalls of this method and future work.

# Preferential attachment model {#sec-gpa}

The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:

Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree where the preference function $b$ is chosen such that:

$$
b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
$$ {#eq-condb1}

$$
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
$$ {#eq-condb2}

Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.

Given conditions [@eq-condb1] and [@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07]. Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:

$$
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
$$
where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.

The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k$, that is:

$$
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
$$
which by the previously stated theorem can also be written as:

$$
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$
Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:

$$
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
$$
We are specifically interested in how the tail heaviness ($\xi$ in the IGPD) is affected by the preference function $b$, [@shimura12] introduces a quantity that will help in determining what domain of attraction this discrete distribution belongs to, that is:

For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let:

$$
\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
$$

[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(k) \sim k^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light tailed. This allows us to show the following:

:::{#prp-omega}
If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then
$$
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$
and

$$
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
$$
:::
See [appendix] for the details of the proof.

@prp-omega aligns with the result from [@krapivsky01] demonstrating that a sub-linear preference function will lead to a light tailed distribution. So, in order for the degree distribution to be heavy tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive. We can in fact show that BA model produces a heavy tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$,  noting that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ is simply $\lim_{k\rightarrow\infty} 1 = 1$ leaving the tail heaviness to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.

We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e. $\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.

Consider the definition of heavy tails:

\begin{align*}
\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
\end{align*}

by Stolz-Cesaro Theorem:

$$
\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
$$

Since $b$ must be asymptotically linear in order to have positive tail heaviness we propose the following preference function:

$$
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.

This preference function, as per @prp-omega, will produce a degree distribution with tail heaviness $\beta/\lambda^*$ guaranteeing a heavy tail. We study this preference function further in the next section.

# Preferential Attachment with flexible heavy tail {#sec-model}

Using a preference function with guaranteed linear behaviour in the limit, allows for the inclusion of sub/super linear behaviour without losing the heavy tails or ending up with a degenerate degree distribution.

The limiting degree distribution resulting from using a preference function of this form can be found to have survival:

$$
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
\end{cases}
$${#eq-polysurv}

with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where:

$$
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
$$

which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.

Some examples of what the degree distribution looks like are shown below in @fig-polylinsurv:

```{r, echo=FALSE}
source('scripts/funcs.R')
```

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical survival distributions of the limiting degree distributions, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
#| label: fig-polylinsurv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = 0:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
}
pars = cbind(pars,surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

ggplot(data = pars) + geom_line(aes(x=(x+1),y=surv, linetype=as.character(b),colour =as.character(b))) +
  scale_x_log10() + scale_y_log10(limits=c(1e-5,1))+theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) + 
  facet_grid(eps~a,labeller = labeller, scales='free')
```

The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Stirling's approximation of the gamma function to obtain:

$$
\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
$$
and utilising the expression from @eq-polysurv:

\begin{align*}
\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
\end{align*}

$$
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
$$ meaning that for $k\ge k_0$ the limiting degree distribution is similar to $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.

To assess how close of an approximation this is the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey. The approximation seems to hold up fairly well even for large degrees.

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical conditional survivals (grey) alongside their IGPD approximations (coloured).
#| label: fig-approx_surv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = k0s:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
gp_surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])/
    S(pars$k0[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
  gp_surv[i] = mev::pgp(pars$x[i],pars$k0[i]-1,((pars$k0[i]+1)^pars$a[i] +pars$eps[i]-1 )/pars$lambda[i],pars$b[i]/pars$lambda[i],lower.tail = F)
}
pars = cbind(pars,surv,gp_surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv','gp_surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

library(ggh4x)

l1 = scale_y_log10(limits=c(1e-1,1))
l2 = scale_y_log10(limits=c(1e-2,1))
l3 = scale_y_log10(limits=c(1e-3,1))
l4 = scale_y_log10(limits=c(1e-4,1))

limlist = list(l1,l2,l3,l4)

ggplot(data = pars) + geom_line(aes(x=(x),y=surv, linetype=as.character(b),colour =as.character(b)), lwd=1) +
  geom_line(data = pars[pars$x>=pars$k0,],aes(x=(x+1),y=gp_surv, linetype=as.character(b)),colour ='grey')+
  scale_x_log10()+
  theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) +
  facet_grid(eps~a,labeller = labeller,scales='free') + facetted_pos_scales(y=limlist)

```

Since $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed. Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices:

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 12
#| fig-height: 6
#| fig-cap: Heat maps of $\xi$ for various combinations of the parameters of the proposed model.
#| label: fig-polyheat
#| 
rho_optim_ba= Vectorize(function(a,eps, b, k0){
  return(abs(rho(2*b, polylin(a, eps), b, k0)-1))
},vectorize.args = 'a')
find_a_ba = Vectorize(function(b, eps, k0){
  out = optimise(rho_optim_ba, c(0.00001,3), b=b, eps=eps, k0=k0)$minimum
  return(out)
}, vectorize.args = 'b')
N =50
as = seq(0,2,l=N+1)[-1]
bs = seq(0,2,l=N+1)[-1]
eps = c(.01,.1,.5,1)
k0 = c(25,100)
pars = expand.grid(as,bs,eps,k0)
names(pars) = c('a', 'b', 'eps', 'k0')
lambdas = numeric(nrow(pars))
a_for_ba = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
  a_for_ba[i] = find_a_ba(pars$b[i], pars$eps[i], pars$k0[i])
}
pars = cbind(pars,lambdas,a_for_ba)
names(pars) = c('a', 'b', 'eps', 'k0','lambda','ba')
labeller = label_bquote(cols = `epsilon`==.(eps),rows = ~k[0]==.(k0))

ggplot(pars) + geom_raster(aes(x=a,y=b,fill=b/lambda)) +
  geom_line(aes(x=ba, y=b),linetype='dashed', colour='red',lwd=1)+
  scale_fill_paletteer_c(palette='grDevices::Blues',limits=c(0,1),direction=-1)+theme(aspect.ratio = 1)+ylim(min(bs), max(bs))+xlim(min(as), max(as))+
  facet_grid(k0~eps,labeller = labeller) + labs(fill=TeX('\\xi')) + xlab(TeX('\\alpha')) + ylab(TeX('\\beta')) + theme_minimal()

```

The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.

For a given network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$ the likelihood is given by:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}

where $B(y,z)$ is the the beta function and $l\ge0$ variable that allows truncating the data such that the minimum degree is $l$.

This likelihood allows for the parameters to be inferred using either a frequentist or a Bayesian approach.

# Recovery {#sec-rec}

The goal of was to provide a possible preference function that is able to explain the growth of real networks with varied tail behaviour in their degree distributions. This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.

The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongisde the priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

to obtain a posterior distribution that can then be used in an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm  to obtain posterior samples. The results of this inference are shown in @fig-rec1 and @fig-rec2. For these simulated networks $l=0$.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
#| label: fig-rec1

surfit_list = list()
pars = readRDS('results/recovery_pars.rds')
recovery_list = readRDS('results/recovery_dat.rds')
selected = c(1,27,14,35)
# selected = 1:36
for(k in 1:length(selected)){
  j=selected[k]
  x = recovery_list[[j]]$mcmc$dat[,1]
  ls = c()

  y_975 = recovery_list[[j]]$mcmc$surv$CI[2,][order(x)]
  y_025 = recovery_list[[j]]$mcmc$surv$CI[1,][order(x)]
  y_50 = recovery_list[[j]]$mcmc$surv$est[order(x)]
  x = sort(x)
  surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
    geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
    scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
    ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
                       ', $\\epsilon = $',pars$eps[j],
                       ', $\\beta = $',pars$b[j])))
}

fig = ggpubr::ggarrange(plotlist = surfit_list,
          nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
          label.x='Degree', label.y = 'Survival')

ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')

```

![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2}

@fig-rec1 and @fig-rec2 demonstrate that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network. This indicates that the method may also be able to be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.

# Application to Real Data {#sec-real}

Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. [describe networks being used]. Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-real1

plt1 = readRDS('results/comp_plot.rds')

plot(plt1)

```

@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model. In most cases, the GPA model does not necessarily provide an improvement in fit when compared to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme. @fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model. Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution and where it doesn't it appears to either be because it is fitting better at the tail than the Zipf-IGP model or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA. This again shows the effects that small degrees have on this model, which is somewhat expected as the theory used for this model is for trees and none of these real networks (nor many real networks) are.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-shapes

plt2 = readRDS('results/shape_plot.rds')

plot(plt2)

```

@fig-pa shows the estimated preference function $b$ alongside the 95% credible interval  on a log-log plot. Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, for those that do the credible interval is much narrower as is the case for `sx-mathoverflow`. The most insightful conclusion we can draw from these plots come from the shape of the preference function. There appears to be two distinct shapes of preference function. The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`. The second distinct shape appears provide some clear PA behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimate for preference function (solid) with 95% credible interval (dashed) on log-log scale.
#| label: fig-pa

plt3 = readRDS('results/PA_plot.rds')
plot(plt3)

```

![: Posterior estimates of paramters for real data.](images/pars_plot.png)

# Conclusion and Discussion {#sec-conc}

In this paper we introduced a flexible class of preference function that when used under the GPA scheme (in the tree setting) is guaranteed to generate a network with a heavy tailed degree distribution whilst remaining flexible in the body. Using simulations from networks using this class of preference function we showed that the model parameters are quite easily estimated from the degrees alone at a snapshot in time. Seeing this we applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way. Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it came with the added benefit of giving a posterior estimate for a preference function.

Obviously this method had its flaws in that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks. Hopefully, as the field progresses we will be able to apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.

GPA is also a fairly simplistic one on its own despite having large flexibility in the preference function allowing for edges to only be added at times when a new node is introduced into the network and not allowing for anything like vertex/edge death, reciprocal edges and rewiring steps. [@deijfen2015] introduces some theory for GPA with vertex death based on the work by [@rudas07], but is unable to come to an expression for the degree distribution. However, adding something like fitness scores into the GPA model seems fairly simple to do even going so far as making each of the parameters used in the class of functions here node-wise instead of universal. For example having the preference weight for a vertex be given by $b(k_i, \underline{\gamma_i})$  where $\underline{\gamma_i}\sim G$ is the vertex's fitness and $k_i$ is the vertex's degree, it is reasonable to expect that the degree distribution for a model like this to be given by:

$$
\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
$$

If this were the case, the posterior distributions of parameters from @sec-real could be interpreted as the approximate distributions from which a vertex's fitness values are drawn instead of providing estimates of the universal parameter values. Additionally, it allows the ability to set some parameters as universal and some as vertex-wise for example one could let $\alpha, \beta, k_0$ be the same for all vertices but let each vertex in the network have its own "zero appeal" $\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)$; moving the model to being more realistic and accounting for the fact that in many real networks the objects vertices represent have some value intrinsic to themselves other than the number of connections they have.

# Appendix {.appendix}

## Additional Results 
### Tail heaviness of general preferential attachment model

Recall the limiting survival function:

$$
\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
$$

The aim is to determine how this distribution behaves for different choices of $b$, specifically what maximum domain of attraction does this distribution belong to and is it affected by the choice of preference function $b$. @shimura12 introduces a quantity that will help in determining what domain of attraction a discrete distribution belongs to, that is:

For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:

$$
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
$$

[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.

Consider a preference function $b(\cdot)$ that fulfills:

$$
\lim_{n\rightarrow \infty} b(n) = \infty
$${#eq-condb1}

Substituting in the form of $\bar F(n)$ from @eq-surv and taking the limit, subject to @eq-condb1:

$$
\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
$$ {#eq-omegares} (see appendix for full proof).

Whilst there are many classes of functions that satisfy @eq-condb1 and @eq-condb2, the vast majority of these result in $\Omega(F,n) \rightarrow 0$, meaning that the limiting degree distribution is light tailed. In fact, in order for the limiting degree distribution to be heavy tailed, $b$ must be asymptotically linear ($b(k)\sim k$).

## Proofs

### Tail heaviness of GPA

Taking the form of the GPA degree survival: $$
\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
$$ and substituting into the formula for $\Omega(F,n)$:

```{=tex}
\begin{align*}
\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
\end{align*}
```
Clearly if $b(n)=c$ or $\lim_{n\rightarrow\infty}b(n)=c$ for some $c>0$ then $\Omega(F,n)=0$. Now consider a non-constant $b(n)$ and re-write $\Omega(F,n)$ as:

```{=tex}
\begin{align*}
\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
\end{align*}
```
Then if $\lim_{n\rightarrow\infty}b(n)=\infty$ it follows that:

```{=tex}
\begin{align*}
\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
\end{align*}
```

### Removing the infinite sum

For a preference function of the form:

$$
b(k) = \begin{cases}
g(k),&k<k_0\\
g(k_0) + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\beta>0, k_0\in\mathbb N$ we have that

```{=tex}
\begin{align*}
\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
\end{align*}
```
Now using the fact that:

$$
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
$$ and reindexing the product in the second sum

```{=tex}
\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
\end{align*}
```
In order to simplify the infinite sum, consider:

```{=tex}
\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
\end{align*}
```
this infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:

```{=tex}
\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
\end{align*}
```

# References
<!-- the pattern below controls the placement of the references -->
:::{#refs}
:::