### 2025-09-22 Files/presentation/pres.qmd

---
title: diaquarto 
author: Author Name
date: "01.11.2024"
subtitle: easy slides in typst
toc: true
format:
  diaquarto-typst: 
    layout: medium # small, medium, large
    ratio: 16/9 # any ratio possible 
    title-color: "154C79" # Hex code for the title color (without #)
---

# About *diatypst*

## Usage

To use this template, add it to your Quarto project with the following command:

```bash
quarto add skriptum/diatypst/diaquarto
```

Then, create a qmd file with the following YAML frontmatter:

```yaml
title: Untitled # required
author: Author Name
date: 01.07.2024
subtitle: easy slides in typst
format:
  diaquarto-typst: # required
    layout: medium 
    ratio: 16/9 
```

Then, insert your content.

- Level-one headings corresponds to new sections.
- Level-two headings corresponds to new slides.
- `---` is used to separate slides.

## Styling

This is the second slide, we test some more content here

> A blockquote

A table:

| A | B | C |
|---|---|---|
| 1 | 2 | 3 |

And finally a

Term
: with a Definition